# HW 1
## 1
See output in appendix. Each row (not including headers) represents a possible combination. ??

a. {(A, U), (A, E), (B, U), (B, E), (C, U), (C, E), (D, U), (D, E), (F, U), (F, E)}

## 2
a. $\mu_{U_1} = 3 + 4 = 7$
<br> $\sigma_{U_1} = 4 * 8 = 32$
b. $\mu_{U_2} = -10 + 4 = -6$
<br> $\sigma_{U_2} = 2 * 8 = 16$
c. $\mu_{U_3} = 1/4 + 4 = 4.25$
<br> $\sigma_{U_3} = 1 * 8 = 8$
d. $\mu_{U_4} = 3/4 + 4 = 4.75$
<br> $\sigma_{U_4} = 1/4 * 8 = 2$

## 3
a. $E[\bar{Y}] = 20$
<br> $\sigma^2[\bar{Y}] = 25 / 10 = 2.5$
b. $E[sum(Y_i)] = 20 * 10 = 200$
<br> $\sigma^2[sum(Y_i)] = 25 * 10 = 250$
c. $E[Y^*] = a + E[\bar{Y}] = a + 20$
<br> $\sigma^2[Y^*] = b^2 * \sigma^2[\bar{Y}] = b^2 * 200$
d. $E[Y^*] = 5 - 2 * 200 = -395$
<br> $\sigma^2[Y^*] = 2^2 * 250 = 1000$

## 4
a. This is a normal distribution because the sum of two normal distributions, $\bar{Y_1}$ and $-\bar{Y_2}$, is also normal.
<br>$E[\bar{Y}_1 - \bar{Y}_2] = \mu_1 - \mu_2$
<br>$\sigma^2[\bar{Y}_1 - \bar{Y}_2] = \sqrt{\frac{\sigma_1^2}{100} - \frac{\sigma_2^2}{100}}$ ??
b. ??

## 5
a. The null hypothesis is that the mean of smokers is equal to the mean of nonsmokers. The alternate hypothesis is that the two means differ.
b. $t_s = 4.4037$ (see appendix for full calculation)
<br> This value is the number of estimated standard deviations our sample mean difference is from the null hypothesis.
c. $p < 0.0005$
d. We reject the null hypothesis because $p < \alpha = 0.05$.

## 6
a. ??

## 7
a. This is a mixed study because there are both observational and experimental variables.
b. The primary variable of interest is test scores.
c. The explanatory variables consists of observational and experimental variables. The high school and section number are the observational variables. The instruction (standard or computer-based) is the experimental variable.
d. See output in appendix. Each row (not including headers) represents a possible combination.

## 8
a. This is a mixed study because there are both observational and experimental variables.
b. The primary variable of interest is number of days required for successful completion.
c. The explanatory variables consists of observational and experimental variables. The prior physical fitness status (below average, average, above average) is the observational variable. The doctor they are paired with (out of three possible doctors) is the experimental variable.
d. See output in appendix. Each row (not including headers) represents a possible combination.

## I
??

# Appendix
```{r}
explanatory_var_combinations <- function(...) {
  # Create a data frame with all combinations
  combinations_df <- expand.grid(...)

  # Convert the data frame to a matrix
  combinations_matrix <- as.matrix(combinations_df)

  # Return the matrix of combinations
  return(combinations_matrix)
}
```

## 1(a)
```{r}
grades <- c("A", "B", "C", "D", "F")
employment <- c("E", "U")

explanatory_var_combinations(grades, employment)
```

## 5(b)
```{r}
mean_1 <- 150.03
mean_2 <- 139.18
s_1 <- 27.49
s_2 <- 27.49
n_1 <- 266
n_2 <- 234

s_p_squared <- (s_1**2 * (n_1 - 1) + s_2**2 * (n_2 - 1)) / (n_1 + n_2 - 2)
t_s <- (mean_1 - mean_2) / sqrt(s_p_squared * (1 / n_1 + 1 / n_2))

t_s
```

## 7(d)
```{r}
high_schools <- c("HS 1", "HS 2", "HS 3")
sections <- c("Section 1", "Section 2", "Section 3", "Section 4")
instructions <- c("computerized", "in person")

explanatory_var_combinations(high_schools, sections, instructions)
```

## 8(d)
```{r}
prior_fitness <- c("fitness below avg", "fitness avg", "fitness above avg")
doctors <- c("doctor 1", "doctor 2", "doctor 3")

explanatory_var_combinations(prior_fitness, doctors)
```

## I-IV
```{r}
data <- read.csv("GSK.csv")
stress <- data$stress
sbp <- data$sbp
exercise <- data$exercise
age <- data$age
education <- data$educatn
marriage <- data$married
gender <- data$gender
```

### I(a)
```{r}
res <- aggregate(sbp ~ stress, data = data, FUN = mean)
colnames(res)[colnames(res) == "sbp"] <- "sbp_mean"

res
```

### I(b)
```{r}
res <- aggregate(sbp ~ stress, data = data, FUN = sd)
colnames(res)[colnames(res) == "sbp"] <- "sbp_sd"

res
```

### I(c)
```{r}
res <- aggregate(age ~ exercise, data = data, FUN = mean)
colnames(res)[colnames(res) == "age"] <- "age_mean"

res
```

### I(d)
```{r}
res <- aggregate(age ~ exercise, data = data, FUN = sd)
colnames(res)[colnames(res) == "age"] <- "age_sd"

res
```

### II(a)
```{r}
boxplot(sbp ~ education, data = data, xlab = "education", ylab = "sbp")
```

### II(b)
```{r}
married <- data[marriage == "Y", "sbp"]
single <- data[marriage == "N", "sbp"]

hist(married, xlab = "sbp")
hist(single, xlab="sbp")
```

### II(c)
```{r}
mosaicplot(table(exercise, stress))
```

### II(d)
```{r}
mosaicplot(table(marriage, stress))
```

### III(a)
```{r}
table(exercise)
```

### III(b)
```{r}
high_ed_stress <- data[education == "HE", "stress"]
table(high_ed_stress)
```

### III(c)
```{r}
res <- aggregate(age ~ stress, data = data, FUN = mean)
colnames(res)[colnames(res) == "age"] <- "age_mean"
res
```

### III(d)
```{r}
res <- aggregate(sbp ~ gender, data = data, FUN = mean)
colnames(res)[colnames(res) == "sbp"] <- "sbp_mean"
res
```

### IV
??