# HW 1
## 1
a. {(A, U), (A, E), (B, U), (B, E), (C, U), (C, E), (D, U), (D, E), (F, U), (F, E)}
<br> this is right! now need to finish ??

## 2
a. $\mu_{U_1} = 3 + 4 = 7$
<br> $\sigma_{U_1} = 4 * 8 = 32$
b. $\mu_{U_2} = -10 + 4 = -6$
<br> $\sigma_{U_2} = 2 * 8 = 16$
c. $\mu_{U_3} = 1/4 + 4 = 4.25$
<br> $\sigma_{U_3} = 1 * 8 = 8$
d. $\mu_{U_4} = 3/4 + 4 = 4.75$
<br> $\sigma_{U_4} = 1/4 * 8 = 2$

## 3
a. $E[\bar{Y}] = 20$
<br> $\sigma^2[\bar{Y}] = 25 / 10 = 2.5$
b. $E[sum(Y_i)] = 20 * 10 = 200$
<br> $\sigma^2[sum(Y_i)] = 25 * 10 = 250$ ??
c. now need to finish ??

## 4
a. This is a normal distribution because the sum of two normal distributions, $\bar{Y_1}$ and $-\bar{Y_2}$, is also normal.
<br>$E[\bar{Y}_1 - \bar{Y}_2] = \mu_1 - \mu_2$
<br>$\sigma^2[\bar{Y}_1 - \bar{Y}_2] = \sqrt{\frac{\sigma_1^2}{100} - \frac{\sigma_2^2}{100}}$ ??
b. ??

## 5
a. The null hypothesis is that the mean of smokers differs from nonsmokers. The alternate hypothesis is that the two means are equal.
b. $t_s = 4.4037$ (see appendix for full calculation)
<br> This value is the number of estimated standard deviations our sample mean difference is from the null hypothesis.
c. ??

## 6
??

## 7
a. This is an experimental study because researchers were modifying the instruction (standard or computer-based) to see if it affects a standardized math test scores.
b. The primary variable of interest is test scores.
c. The explanatory variable is the instruction.
d. complete ??

## 8
??

## I
??

# Appendix
## 5b
```{r}
mean_1 <- 150.03
mean_2 <- 139.18
s_1 <- 27.49
s_2 <- 27.49
n_1 <- 266
n_2 <- 234

s_p_squared <- (s_1**2 * (n_1 - 1) + s_2**2 * (n_2 - 1)) / (n_1 + n_2 - 2)
t_s <- (mean_1 - mean_2) / sqrt(s_p_squared * (1 / n_1 + 1 / n_2))

t_s
```

## I-IV
```{r}
data <- read.csv("GSK.csv")
stress <- data$stress
sbp <- data$sbp
exercise <- data$exercise
age <- data$age
education <- data$educatn
marriage <- data$married
gender <- data$gender
```

### I(a)
```{r}
res <- aggregate(sbp ~ stress, data = data, FUN = mean)
colnames(res)[colnames(res) == "sbp"] <- "sbp_mean"

res
```

### I(b)
```{r}
res <- aggregate(sbp ~ stress, data = data, FUN = sd)
colnames(res)[colnames(res) == "sbp"] <- "sbp_sd"

res
```

### I(c)
```{r}
res <- aggregate(age ~ exercise, data = data, FUN = mean)
colnames(res)[colnames(res) == "age"] <- "age_mean"

res
```

### I(d)
```{r}
res <- aggregate(age ~ exercise, data = data, FUN = sd)
colnames(res)[colnames(res) == "age"] <- "age_sd"

res
```

### II(a)
```{r}
boxplot(sbp ~ education, data = data, xlab = "education", ylab = "sbp")
```

### II(b)
```{r}
married <- data[marriage == "Y", "sbp"]
single <- data[marriage == "N", "sbp"]

hist(married, xlab = "sbp")
hist(single, xlab="sbp")
```

### II(c)
```{r}
mosaicplot(table(exercise, stress))
```

### II(d)
```{r}
mosaicplot(table(marriage, stress))
```

### III(a)
```{r}
table(exercise)
```

### III(b)
```{r}
high_ed_stress <- data[education == "HE", "stress"]
table(high_ed_stress)
```

### III(c)
```{r}
res <- aggregate(age ~ stress, data = data, FUN = mean)
colnames(res)[colnames(res) == "age"] <- "age_mean"
res
```

### III(d)
```{r}
res <- aggregate(sbp ~ gender, data = data, FUN = mean)
colnames(res)[colnames(res) == "sbp"] <- "sbp_mean"
res
```

### IV
??