```{r echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# 1
a. $\hat{\mu}_1 = 38$, $\hat{\mu}_2 = 32$, $\hat{\mu}_3 = 24$
b. $E[Y_{1j}] = 38$, $E[Y_{2j}] = 32$, $E[Y_{3j}] = 24$
c. $SSTO = s^2_Y (n_T - 1) = 1088.691$
<br> $SSA = 672$
<br> $SSE = SSTO - SSA = 416.6912$
```{r}

# 1-c
s <- 6.88
n <- 24
n_i <- c(8, 10, 6)

ssto <- s ** 2 * (n - 1)

y_bar_i_dot <- c(38, 32, 24)
y_bar_dot_dot <- 32

ssa <- sum(n_i * (y_bar_i_dot - y_bar_dot_dot) ** 2)
sse <- ssto - ssa

ssto
ssa
sse
```
d. $d.f.[SSTO] = n_T - 1 = 23$
<br> $d.f.[SSA] = a - 1 = 2$
<br> $d.f.[SSE] = n_T - a = 21$
```{r}

# 1-d
a <- 3

df_ssto <- n - 1
df_ssa <- a - 1
df_sse <- n - a

df_ssto
df_ssa
df_sse
```
e. $MSTO = \frac{SSTO}{d.f.[SSTO]} = 47.3344$
<br> $MSA = \frac{SSA}{d.f.[SSA]} = 336$
<br> $MSE = \frac{SSE}{d.f.[SSE]} = 19.84244$
```{r}

# 1-e
ssto / df_ssto
ssa / df_ssa
sse / df_sse
```

# 2
a. The variance for each group is not constant because the variances between groups are different.
b. Seems like group 3 has the fastest recovery time. We cannot tell if it's statistically significantly faster because the mean of group 3 is within 3 standard deviations of the slowest group. ??

# Appendix
```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```