```{r echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# 1
```{r}

# 1-a
mu_T <- 32
mu_i <- c(38, 32, 24)
n_i <- c(8, 10, 6)
a <- 3
MSE <- 19.8424 # From HW2

phi <- round(1 / sqrt(MSE) * sqrt(sum(n_i * (mu_i - mu_T) ** 2) / a), 4)
```
a. $\phi =$ `r phi`
b. Using $\phi = `r phi`$, $d.f.\{SSA\} = 2$, and $d.f.\{SSE\} = 21$ from HW2, we use the table to find that the power is $0.99$.
```{r}

# 1-c
mu_min <- min(mu_i)
mu_max <- max(mu_i)
effect_size <- round((mu_max - mu_min) / sqrt(MSE), 4)
```
c. $\frac{\hat{\Delta}}{\sigma_\epsilon} = `r effect_size`$
d. Using the table, we find that the minimum sample size for each group is $4$.

# 2
a. For all $Y_{ij} = \mu_i + \epsilon_{ij}$, we are interested to know the factor effect ($\gamma_i$) and confidence interval ($\mu_i - \mu_.$) of each group. The assumptions are:
    i. $Y_{ij}$ were randomly sampled (independent)
    ii. The $i$ groups are independent for all $i = 1, ..., a$
    iii. Errors are independent and $\epsilon_{ij} \sim N(0, \sigma_\epsilon^2)$
```{r}

# 2-b
Y_bar_bar <- sum(mu_i) / a
gamma_i <- round(mu_i - Y_bar_bar, 4)
```
b. $\gamma_1 = `r gamma_i[1]`, \gamma_2 = `r gamma_i[2]`, \gamma_3 = `r gamma_i[3]`$ 
c. $\gamma_2$ is the factor effect, or how the overall mean is effected by a subject being in group $2$.
```{r}

# 2-d
gamma_T <- round(sum(gamma_i))
```
d. $\sum \gamma_i = `r gamma_T`$
```{r}

# 2-e
sd_Y_i <- c(5.48, 3.46, 4.43)
var_Y_i <- sd_Y_i ** 2
var_Y_bar_bar <- round(sum(var_Y_i / n_i) / a ** 2, 4)
```
e. If the ANOVA model holds, the distribution of $\hat{\gamma}_k$ is normal where $\hat{\gamma}_k \sim N(\gamma_k, \sigma^2[\hat{\gamma}_k])$.
<br> $\sigma^2[\hat{\gamma}_k] = \sigma^2[\bar{Y}_{k.} - \bar{\bar{Y}}]$
<br> $\to \sigma^2[\bar{Y}_{k.}] + \sigma^2[\bar{\bar{Y}}]$
<br> $\to \sigma^2[\bar{Y}_{k.}] + \sigma^2[\frac{\sum{\bar{Y}_{i.}}}{a}]$
<br> $\to \sigma^2[\bar{Y}_{k.}] + \sigma^2[\frac{\sum{\bar{Y}_{i.}}}{`r a`}]$
<br> $\to \sigma^2[\bar{Y}_{k.}] + \frac{\sum{\sigma^2[\bar{Y}_{i.}]}}{`r a ** 2`}$
<br> $\to \frac{\sigma^2[Y_k]}{n_k} + \frac{1}{`r a ** 2`} \sum{\frac{\sigma^2[Y_i]}{n_i}}$
<br> $\to \frac{\sigma^2[Y_k]}{n_k} + `r var_Y_bar_bar`$
<br> $\therefore \hat{\gamma}_k \sim N(\gamma_k, \frac{\sigma^2[Y_k]}{n_k} + `r var_Y_bar_bar`)$ ??

# 3
```{r}

# 3-a
alpha <- 0.05
n_T <- 24
a <- 3
t_critical <- qt(1 - alpha / 2, n_T - a)
lower <- round(mu_i[1] - t_critical * sqrt(MSE * (1 / n_i[1])), 4)
upper <- round(mu_i[1] + t_critical * sqrt(MSE * (1 / n_i[1])), 4)
```
a. The $95%$ confidence interval for $\mu_1$ is $[`r lower`, `r upper`]$. This means that we are 95% sure that the true value of $\mu_1 \in [`r lower`, `r upper`]$.
```{r}

# 3-b
alpha <- 0.01
n_T <- sum(n_i)
critical_t <- qt(1 - alpha / 2, n_T - a)
lower_bound <- round(
  mu_i[1] - mu_i[3] - critical_t * sqrt(MSE * (1 / n_i[1] + 1 / n_i[3])),
  4
)
upper_bound <- round(
  mu_i[1] - mu_i[3] + critical_t * sqrt(MSE * (1 / n_i[1] + 1 / n_i[3])),
  4
)
```
b. $\mu_1 - \mu_3 \in [`r lower_bound`, `r upper_bound`]$
<br> $\mu_1$ is between $`r lower_bound`$ and $`r upper_bound`$ more than $\mu_3$. Since all values in the confidence interval is positive, we believe that the two means are different.
c. Yes, my confidence interval agrees with that statement. We expect the recovery time for "Below Average" patient to be at least $`r lower`$ days, which is greater than $30$ days.
d. The largest difference we expect an above average patient to have compared to a below average in average recovery time is $`r upper_bound`$ as it is our upper bound of our confidence interval for the difference between the two means. 

# 4
a. The power is the probability that we correctly reject the null hypothesis where age does affect the cash offer.
```{r}

# 4-b
n_i <- c(12, 12, 12)
mu_i <- c(27.75, 21.42, 21.50)
mu_T <- 23.56
a <- 3
MSE <- 2.5243 # from HW2

phi <- round(1 / sqrt(MSE) * sqrt(sum(n_i * (mu_i - mu_T) ** 2) / a), 4)
```
b. $\phi = `r phi`$
c. Using $\phi = `r phi`$, $d.f.\{SSA\} = 2$, and $d.f.\{SSE\} = 33$ from HW2, we use the table to find that the power is $1.00$.
d. The power is super high because our factor effect is high relative to the error (i.e. $SSA$ is a lot higher than $MSE$).

# 5
a. For all $Y_{ij} = \mu_i + \epsilon_{ij}$, we are interested to know the factor effect ($\gamma_i$) and confidence interval ($\mu_i - \mu_.$) of each group. The assumptions are:
    i. $Y_{ij}$ were randomly sampled (independent)
    ii. The $i$ groups are independent for all $i = 1, ..., a$
    iii. Errors are independent and $\epsilon_{ij} \sim N(0, \sigma_\epsilon^2)$
```{r}

# 5-a
M <- 1
Y <- 3
O <- 2
mu_diff <- mu_i[M] - mu_i[Y]
```
b. $\mu_M - \mu_Y = `r mu_diff`$
c. $\sigma^2[\bar{M} - \bar{Y}]$
<br> $\to \sigma^2[\bar{M}] + \sigma^2[\bar{Y}]$ ??
```{r}

# 5-d
sd <- round(sqrt(MSE), 4)
sd_above_mean <- round(2.5 / sqrt(MSE), 4)
```
d. $\epsilon_{ij} \sim N(0, \sqrt{MSE})$
<br> $\to \epsilon_{ij} \sim N(0, `r sd`)$
<br> This means that $2.5$ is $`r sd_above_mean`$ standard deviations above the mean. Using the z-table, $p \in [0.0571, 0.0582]$.

# 6
```{r}

# 6-a
n_T <- sum(n_i)
alpha <- 0.05
CI_mid <- mu_i[M] - (mu_i[Y] + mu_i[O]) / 2
t_critical <- qt(1 - alpha / 2, n_T - a)
lower <- round(
  CI_mid - t_critical * sqrt(
    MSE * (1 / n_i[M] + 1 / 2 ** 2 * 1 / n_i[Y] + 1 / 2 ** 2 * 1 / n_i[O])
  ),
  4
)
upper <- round(
  CI_mid + t_critical * sqrt(
    MSE * (1 / n_i[M] + 1 / 2 ** 2 * 1 / n_i[Y] + 1 / 2 ** 2 * 1 / n_i[O])
  ),
  4
)
```
a. The $95%$ confidence interval for $\mu_M - \frac{\mu_Y + \mu_O}{2}$ is $[`r lower`, `r upper`]$.
b. We are $95%$ confident that $\mu_M - \frac{\mu_Y + \mu_O}{2} \in [`r lower`, `r upper`]$.
```{r}

# 6-c
CI_mid <- mu_i[O] - mu_i[Y]
lower <- round(
  CI_mid - t_critical * sqrt(
    MSE * (1 / n_i[O] + 1 / n_i[Y])
  ),
  4
)
upper <- round(
  CI_mid + t_critical * sqrt(
    MSE * (1 / n_i[O] + 1 / n_i[Y])
  ),
  4
)
```
c. The $95%$ confidence interval for $\mu_O - \mu_Y$ is $[`r lower`, `r upper`]$.
d. We are $95%$ confident that $\mu_O - \mu_Y \in [`r lower`, `r upper`]$.
e. I believe the means from 6A are significantly different because the confidence interval does not contain zero. The means from 6C are not significantly different because the confidence interval does not contain 0.

## 7
```{r}

# 7-a
mu_i <- c(13, 13, 18, 18)
n_i <- 6
a <- 4
sd_err <- 3.5

n_T <- n_i * a
mu_T <- sum(mu_i * n_i) / n_T
df_num <- a - 1
df_den <- n_T - a

phi <- round(1 / sd_err * sqrt(sum(n_i * (mu_i - mu_T) ** 2) / a), 4)
```
a. $\phi = `r phi`, d.f.[num] = `r df_num`, d.f.[den] = `r df_den`, \alpha = 0.05$.
<br> Using the F-table, our power is $0.62$.
b. $\phi = `r phi`, d.f.[num] = `r df_num`, d.f.[den] = `r df_den`, \alpha = 0.01$.
<br> Using the F-table, our power is $0.34$.
c. As we decrease $\alpha$, the power goes down.
```{r}

# 7-d
delta_over_sd_err <- round((max(mu_i) - min(mu_i)) / sd_err, 4)
```
d. $\frac{\Delta}{\sigma_\epsilon} = `r delta_over_sd_err`, a = `r a`, \alpha = 0.05, 1 - \beta = 0.80$.
<br> Using the sample size table, we find that $n_i = 11$ is our minimum sample size for all groups.
e. $\frac{\Delta}{\sigma_\epsilon} = `r delta_over_sd_err`, a = `r a`, \alpha = 0.05, 1 - \beta = 0.90$.
<br> Using the sample size table, we find that $n_i = 14$ is our minimum sample size for all groups.
f. As we aim for a higher power, we need a larger sample size for each group.

## 8
??

# Appendix
```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```