---
title: "HW 6"
author: "Andrew Jowe"
output:
  html_document:
    df_print: kable
  pdf_document:
    df_print: kable
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r Functions}
# Plus minus function
plus.minus = function(value) {
  return(c(-value, value))
}

# Give me multipliers
find.mult = function(alpha,a,b,dfSSE,g,group){
  if(group == "A"){
    Tuk = round(qtukey(1-alpha,a,dfSSE)/sqrt(2),3)
    Bon = round(qt(1-alpha/(2*g), dfSSE ) ,3)
    Sch = round(sqrt((a-1)*qf(1-alpha, a-1, dfSSE)),3)
  }else if(group == "B"){
    Tuk = round(qtukey(1-alpha,b,dfSSE)/sqrt(2),3)
    Bon = round(qt(1-alpha/(2*g), dfSSE ) ,3)
    Sch = round(sqrt((b-1)*qf(1-alpha, b-1, dfSSE)),3)
  }else if(group == "AB"){
    Tuk = round(qtukey(1-alpha,a*b,dfSSE)/sqrt(2),3)
    Bon = round(qt(1-alpha/(2*g), dfSSE ) ,3)
    Sch = round(sqrt((a*b-1)*qf(1-alpha, a*b-1, dfSSE)),3)
  }
  results = c(Bon, Tuk,Sch)
  names(results) = c("Bonferroni","Tukey","Scheffe")
  return(results)
}

give.me.CI = function(the.data,MSE,equal.weights = TRUE,multiplier,group,cs){
   if(sum(cs) != 0 & sum(cs !=0 ) != 1){
    return("Error - you did not input a valid contrast")
  }else{
    the.means = find.means(the.data)
    the.ns =find.means(the.data,length)
    nt = nrow(the.data)
    a = length(unique(the.data[,2]))
    b = length(unique(the.data[,3]))
    if(group =="A"){
      if(equal.weights == TRUE){
        a.means = rowMeans(the.means$AB)
        est = sum(a.means*cs)
        mul = rowSums(1/the.ns$AB)
        SE = sqrt(MSE/b^2 * (sum(cs^2*mul)))
        N = names(a.means)[cs!=0]
        CS = paste("(",cs[cs!=0],")",sep = "")
        fancy = paste(paste(CS,N,sep =""),collapse = "+")
        names(est) = fancy
      } else{
        a.means = the.means$A
        est = sum(a.means*cs)
        SE = sqrt(MSE*sum(cs^2*(1/the.ns$A)))
        N = names(a.means)[cs!=0]
        CS = paste("(",cs[cs!=0],")",sep = "")
        fancy = paste(paste(CS,N,sep =""),collapse = "+")
        names(est) = fancy
      }
    }else if(group == "B"){
      if(equal.weights == TRUE){
        b.means = colMeans(the.means$AB)
        est = sum(b.means*cs)
        mul = colSums(1/the.ns$AB)
        SE = sqrt(MSE/a^2 * (sum(cs^2*mul)))
        N = names(b.means)[cs!=0]
        CS = paste("(",cs[cs!=0],")",sep = "")
        fancy = paste(paste(CS,N,sep =""),collapse = "+")
        names(est) = fancy
      } else{
        b.means = the.means$B
        est = sum(b.means*cs)
        SE = sqrt(MSE*sum(cs^2*(1/the.ns$B)))
        N = names(b.means)[cs!=0]
        CS = paste("(",cs[cs!=0],")",sep = "")
        fancy = paste(paste(CS,N,sep =""),collapse = "+")
        names(est) = fancy
      }
    } else if(group == "AB"){
      est = sum(cs*the.means$AB)
      SE = sqrt(MSE*sum(cs^2/the.ns$AB))
      names(est) = "someAB"
    }
    the.CI = est + c(-1,1)*multiplier*SE
    results = c(est,the.CI)
    names(results) = c(names(est),"lower bound","upper bound")
    return(results)
  }
}
```

# 1
a. The overall trend for factor A is that as we increase the dosage of the first ingredient, the number of hours of relief increases.
b. The overall trend for factor B is that as we increase the dosage of the second ingredient, the number of hours of relief increases.
c. I would use the smallest multiplier out of the three for precision since all multipliers can be used.
d. If we assume no interaction effect, I would say increasing factor A is generally more important for increasing the number of hours of relief since there is a bigger increase from low to medium and medium to high for factor A compared to factor B.

# 2

```{r 2.a}
a <- 3
b <- 3
nidot <- 12
ndotj <- 12
nT <- 36

alpha <- 0.05
g <- 2

SSE <- 1.63
dfSSE <- nT - a * b
MSE <- SSE / dfSSE

mu1dot <- 3.88
mudot1 <- 4.63

all.mult <- find.mult(alpha = alpha, a = a, b = b,
                      dfSSE = dfSSE, g = g, group = "AB")
the.mult <- min(all.mult)
the.CI <- mu1dot - mudot1 + plus.minus(
  the.mult * sqrt(MSE * (1 / nidot + 1 / ndotj))
)
```

a. The 95% confidence interval is $[`r the.CI[1]`, `r the.CI[2]`]$.
b. We are 95% certain that the difference of hours of relief on average between low dosage of factor A and low dosage of factor B is around `r min(abs(the.CI))` to `r max(abs(the.CI))` lower. This means for low dosage of either drug, people are better off with drug B.
```{r 2.c}
alpha <- 0.01
g <- 2

mu3dot <- 9.83
mudot3 <- 8.98

all.mult <- find.mult(alpha = alpha, a = a, b = b,
                      dfSSE = dfSSE, g = g, group = "AB")
the.mult <- min(all.mult)
the.CI.1 <- mu1dot - mu3dot + plus.minus(
  the.mult * sqrt(MSE * (1 / nidot + 1 / nidot))
)
the.CI.2 <- mudot1 - mudot3 + plus.minus(
  the.mult * sqrt(MSE * (1 / ndotj + 1 / ndotj))
)
```
c. The 99% confidence interval for $\mu_{1.} - \mu_{3.}$ is $[`r the.CI.1[1]`, `r the.CI.1[2]`]$. The 99% confidence interval for $\mu_{.1} - \mu_{.3}$ is $[`r the.CI.2[1]`, `r the.CI.2[2]`]$.
d. The confidence interval for $\mu_{1.} - \mu_{3.}$ tells us that we are 99% certain that difference of hours of relief on average between a low dosage of drug A and a high dosage of drug A is around `r min(abs(the.CI.1))` to `r max(abs(the.CI.1))` lower.
```{r 2.e}
alpha <- 0.1
g <- 1

nij <- 4

mu11 <- 2.48
mu13 <- 4.58

all.mult <- find.mult(alpha = alpha, a = a, b = b,
                      dfSSE = dfSSE, g = g, group = "AB")
the.mult <- min(all.mult)
the.CI <- mu11 - mu13 + plus.minus(
  the.mult * sqrt(MSE * (1 / nij + 1 / nij))
)
```
e. The 90% confidence interval is $[`r the.CI[1]`, `r the.CI[2]`]$. We are 90% certain that the difference of hours of relief on average between low dosage of factor A and high dosage of factor A is around `r min(abs(the.CI))` to `r max(abs(the.CI))` lower for those who took low dosage of factor B.

# 3

a. Yes, because 0 is not in our interval.
b. Yes, because 0 is not in our interval.
c. Yes, because 0 is not in our interval.
d. I would have them take a high dosage of drug A and drug B because we have clearly shown higher dosages of both drugs yields better results.

# 4
a. Yes, a weighted confidence interval changes the bounds compared to the unweighted confidence interval. Lets look at the equation for confidence interval:
    - $\sum\sum c_{ij} \bar{Y_{ij}} \pm t_{\alpha/2, df(SSE)} \sqrt{MSE \sum \sum \frac{c_{ij}^2}{n_{ij}}}$
    - We can tell that the weights can change not only the center of the confidence interval but also the spread of the confidence interval. Both of these will affect the bounds.
b. ??
c. One reason to give unequal weights for smokers is due to the unequal sample sizes. Since the sample size of smokers is larger, we can assign a larger weight to this as it is more reliable information and indicates greater precision.
d. One reason to give equal weights is because unequal weights makes interpretation of the result difficult.

# 5
a. We are 95% certain that females on average have around 0.84 to 2.18 less lung capacity compared to males.
b. It would be unusual because -3 is not in our confidence interval.
c. We are 95% certain that nonsmokers on average have around 1.82 to 2.45 more lung capacity compared to smokers.
d. Yes, because 0 is not in our confidence interval.

# 6
a. Lets interpret the confidence interval for $\mu_{11} - \mu_{12}$. The difference of lung capacity on average between non smokers and smokers is 0.4 to 2.6 more for females.
b. No, because our interval from (a) includes the interaction effect of A on B compared to (5c) which only accounted for factor B.
c. No, because our interval from (a) includes the interaction effect of A on B compared to (5c) which only accounted for factor B.
d. $\beta_0 = 7.30$ is our default modifier regardless of which group the subject belongs to. $\beta_1 = 1.32$ is our modifier if the subject is within Male group of factor A. $\beta_2 = -2.180$ is our modifier if the subject is within the smokers group of factor B.

# 7
a. True. If there is no interaction effect, we don't expect a change in confidence interval if we choose to look at different groups in factor B when contrasting two means in factor A. We might as well compute the confidence interval for a contrast of $\mu_{i.}$ or $\mu_{.j}$.
b. False. The multiplier used should be the smallest multiplier out of all the multipliers. Even though Bonferroni multiplier can be used for any contrast, it is not necessarily the smallest.
c. False. These multipliers are used for creating confidence intervals that contrast means, not for finding the true population mean.
d. True. The amount of parameters to estimate for the regression model is the same as the model we chose. In our case, we chose the no interaction model, which has $a + b - 1$ parameters.

# Appendix
```{r appendix, results="asis"}
# Get all the labels
all_labels <- knitr::all_labels()

# Print out code along with labels
for (label in all_labels) {
  if (label == "setup" || label == "appendix") {
    next
  }

  code <- knitr::knit_code$get(label)

  cat("##", label, "\n")
  cat("```r \n")
  cat(code, sep = "\n")
  cat("```\n")
}
```