---
title: "Project 2 Outliers and Transformation"
author: "Andrew Jowe"
output:
  html_document:
    df_print: kable
  pdf_document:
    df_print: kable
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Assumption
$\alpha = 0.1$ for both const variance and normality test.

# Import dataset and get model fit
```{r 1}
the.data <- read.csv("NewHawk.csv")
the.model <- lm(Wing ~ Species, data = the.data)
model.fit <- aov(Wing ~ Species,
  data = the.data
)
```

# Get QQ Plot
```{r}
car::qqPlot(model.fit$residuals,
  id = FALSE # remove point identification
)
```

# Do test for normality
```{r}
the.SWtest = shapiro.test(model.fit$residuals)
the.SWtest
```

# Do plot for variance
```{r}
plot(model.fit, which = 3)
```

# Do test for const variance
```{r}
car::leveneTest(model.fit)
```

# Transformation Possibilities
1. PPCC
2. Shapiro-Wilks
3. Log-Likelihood 
4. 
```{r}
# All transformations considered
#QQplot
L1 <- EnvStats::boxcox(model.fit, objective.name = "PPCC", optimize = TRUE)$lambda
#Shapiro-Wilks
L2 <- EnvStats::boxcox(model.fit, objective.name = "Shapiro-Wilk", optimize = TRUE)$lambda
L3 <- EnvStats::boxcox(the.data$Wing, objective.name = "Log-Likelihood",optimize = TRUE)$lambda
```

# No outlier removal, PPCC Transformation
```{r}
# The transformation function to get transformed model
give.me.t.model <- function(L, data = the.data) {
  YT = (data$Wing^(L) - 1) / L
  t.data = data.frame(Wing = YT, Species = data$Species)
  t.model = aov(Wing ~ Species, data = t.data)
  return(t.model)
}
```

```{r}
# Get summary of this model
t.model <- give.me.t.model(L1)
summary(t.model)
```

## Get QQ Plot
```{r}
car::qqPlot(t.model$residuals,
id = FALSE # remove point identification
)
```

## Test for normality
```{r}
the.SWtest = shapiro.test(t.model$residuals)
the.SWtest
```

## Plot variances
```{r}
# Homogeneity of variances
plot(t.model, which = 3)
```

## Test for const variance
```{r}
# Levene test
car::leveneTest(t.model)
```

# No outlier removal, Log Likelihood Transformation
```{r}
# Get summary of this model
t.model <- give.me.t.model(L3)
summary(t.model)
```

## Get QQ Plot
```{r}
car::qqPlot(t.model$residuals,
id = FALSE # remove point identification
)
```

## Test for normality
```{r}
the.SWtest = shapiro.test(t.model$residuals)
the.SWtest
```

## Plot variances
```{r}
# Homogeneity of variances
plot(t.model, which = 3)
```

## Test for const variance
```{r}
# Levene test
car::leveneTest(t.model)
```

# No outlier removal, SW Transformation
```{r}
# Get summary of this model
t.model <- give.me.t.model(L2)
summary(t.model)
```

## Get QQ Plot
```{r}
car::qqPlot(t.model$residuals,
id = FALSE # remove point identification
)
```

## Test for normality
```{r}
the.SWtest = shapiro.test(t.model$residuals)
the.SWtest
```

## Plot variances
```{r}
# Homogeneity of variances
plot(t.model, which = 3)
```

## Test for const variance
```{r}
# Levene test
car::leveneTest(t.model)
```

# Possible outlier removal techniques
1. Outlier removal via box plot
2. Semi-Studentized Residuals: we can use this since we have the assumption that our variance is constant from our original test. We don't need to do studentized residuals since this is a more robust replacement.

# Removing outliers via box plot (1)
```{r}
# remove outliers
the.box.plot <- boxplot(Wing ~ Species, data = the.data)
outliers <- the.box.plot$out
data.no.outlier.1 <- the.data[!the.data$Wing %in% outliers, ]
```

```{r}
# Get summary of this model
t.model <- aov(Wing ~ Species, data = data.no.outlier.1)
model.no.outlier.1 <- t.model
summary(t.model)
```

## Get QQ Plot
```{r}
car::qqPlot(t.model$residuals,
  id = FALSE # remove point identification
)
```

## Test for normality
```{r}
the.SWtest = shapiro.test(t.model$residuals)
the.SWtest
```

## Plot variances
```{r}
# Homogeneity of variances
plot(t.model, which = 3)
```

## Test for const variance
```{r}
# Levene test
car::leveneTest(t.model)
```

# Removing outliers via Studentized Residuals (2)
$\alpha = 0.05$

```{r}
rij = rstandard(the.model)
alpha = 0.05

nt = nrow(the.data) #Calculates the total sample size
a = length(unique(the.data$Species)) #Calculates the value of a
t.cutoff= qt(1-alpha/(2*nt), nt-a)
CO.rij = which(abs(rij) > t.cutoff)

outliers = CO.rij
data.no.outlier.2 = the.data[-outliers,]
```

```{r}
# Get summary of this model
t.model <- aov(Wing ~ Species, data = data.no.outlier.2)
model.no.outlier.2 <- t.model
summary(t.model)
```

## Get QQ Plot
```{r}
car::qqPlot(t.model$residuals,
  id = FALSE # remove point identification
)
```

## Test for normality
```{r}
the.SWtest = shapiro.test(t.model$residuals)
the.SWtest
```

## Plot variances
```{r}
# Homogeneity of variances
plot(t.model, which = 3)
```

## Test for const variance
```{r}
# Levene test
car::leveneTest(t.model)
```

# Remove outliers (1) and PPCC Transformation
```{r}
# All transformations considered
#QQplot
L1 <- EnvStats::boxcox(model.no.outlier.1, objective.name = "PPCC", optimize = TRUE)$lambda
#Shapiro-Wilks
L2 <- EnvStats::boxcox(model.no.outlier.1, objective.name = "Shapiro-Wilk", optimize = TRUE)$lambda
L3 <- EnvStats::boxcox(data.no.outlier.1$Wing, objective.name = "Log-Likelihood",optimize = TRUE)$lambda
```

```{r}
# Get summary of this model
t.model <- give.me.t.model(L1, data = data.no.outlier.1)
summary(t.model)
```

## Get QQ Plot
```{r}
car::qqPlot(t.model$residuals,
id = FALSE # remove point identification
)
```

## Test for normality
```{r}
the.SWtest = shapiro.test(t.model$residuals)
the.SWtest
```

## Plot variances
```{r}
# Homogeneity of variances
plot(t.model, which = 3)
```

## Test for const variance
```{r}
# Levene test
car::leveneTest(t.model)
```

# Remove outliers (1), SW Transformation
```{r}
# Get summary of this model
t.model <- give.me.t.model(L2, data = data.no.outlier.1)
summary(t.model)
```

## Get QQ Plot
```{r}
car::qqPlot(t.model$residuals,
id = FALSE # remove point identification
)
```

## Test for normality
```{r}
the.SWtest = shapiro.test(t.model$residuals)
the.SWtest
```

## Plot variances
```{r}
# Homogeneity of variances
plot(t.model, which = 3)
```

## Test for const variance
```{r}
# Levene test
car::leveneTest(t.model)
```

# Remove outliers (1), Log Likelihood Transformation
```{r}
# Get summary of this model
t.model <- give.me.t.model(L3, data = data.no.outlier.1)
summary(t.model)
```

## Get QQ Plot
```{r}
car::qqPlot(t.model$residuals,
id = FALSE # remove point identification
)
```

## Test for normality
```{r}
the.SWtest = shapiro.test(t.model$residuals)
the.SWtest
```

## Plot variances
```{r}
# Homogeneity of variances
plot(t.model, which = 3)
```

## Test for const variance
```{r}
# Levene test
car::leveneTest(t.model)
```

# Remove outliers (2) and PPCC Transformation
```{r}
# All transformations considered
#QQplot
L1 <- EnvStats::boxcox(model.no.outlier.2, objective.name = "PPCC", optimize = TRUE)$lambda
#Shapiro-Wilks
L2 <- EnvStats::boxcox(model.no.outlier.2, objective.name = "Shapiro-Wilk", optimize = TRUE)$lambda
L3 <- EnvStats::boxcox(data.no.outlier.2$Wing, objective.name = "Log-Likelihood",optimize = TRUE)$lambda
```

```{r}
# Get summary of this model
t.model <- give.me.t.model(L1, data = data.no.outlier.2)
summary(t.model)
```

## Get QQ Plot
```{r}
car::qqPlot(t.model$residuals,
id = FALSE # remove point identification
)
```

## Test for normality
```{r}
the.SWtest = shapiro.test(t.model$residuals)
the.SWtest
```

## Plot variances
```{r}
# Homogeneity of variances
plot(t.model, which = 3)
```

## Test for const variance
```{r}
# Levene test
car::leveneTest(t.model)
```

# Remove outliers (2), SW Transformation
```{r}
# Get summary of this model
t.model <- give.me.t.model(L2, data = data.no.outlier.2)
summary(t.model)
```

## Get QQ Plot
```{r}
car::qqPlot(t.model$residuals,
id = FALSE # remove point identification
)
```

## Test for normality
```{r}
the.SWtest = shapiro.test(t.model$residuals)
the.SWtest
```

## Plot variances
```{r}
# Homogeneity of variances
plot(t.model, which = 3)
```

## Test for const variance
```{r}
# Levene test
car::leveneTest(t.model)
```

# Remove outliers (2), Log Likelihood Transformation
```{r}
# Get summary of this model
t.model <- give.me.t.model(L3, data = data.no.outlier.2)
summary(t.model)
```

## Get QQ Plot
```{r}
car::qqPlot(t.model$residuals,
id = FALSE # remove point identification
)
```

## Test for normality
```{r}
the.SWtest = shapiro.test(t.model$residuals)
the.SWtest
```

## Plot variances
```{r}
# Homogeneity of variances
plot(t.model, which = 3)
```

## Test for const variance
```{r}
# Levene test
car::leveneTest(t.model)
```

# Conclusion
There are no good combination of transformed variables, since for any combination of outlier removal and transformation, we will result with either non-normal data or unequal variance or both.