---
title: "Analysis"
author: "Andrew"
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r Setup}
alpha <- 0.05

data <- read.csv("insurance.csv")
selected_columns <- c("charges", "children", "region")
data <- subset(data, select = selected_columns)
colnames(data) <- c("Y", "A", "B")

AB <- lm(Y ~ A * B, data = data)
A.B <- lm(Y ~ A + B, data = data)
A <- lm(Y ~ A, data = data)
B <- lm(Y ~ B, data = data)
N <- lm(Y ~ 1, data = data)
```

```{r Functions}
# Give me means
find.means = function(data,fun.name = mean){
  a = length(unique(data[,2]))
  b = length(unique(data[,3]))
  means.A = by(data[,1], data[,2], fun.name)
  means.B = by(data[,1],data[,3],fun.name)
  means.AB = by(data[,1],list(data[,2],data[,3]),fun.name)
  MAB = matrix(means.AB,nrow = b, ncol = a, byrow = TRUE)
  colnames(MAB) = names(means.A)
  rownames(MAB) = names(means.B)
  MA = as.numeric(means.A)
  names(MA) = names(means.A)
  MB = as.numeric(means.B)
  names(MB) = names(means.B)
  MAB = t(MAB)
  results = list(A = MA, B = MB, AB = MAB)
  return(results)
}

# Partial R^2
Partial.R2 = function(small.model,big.model){
  SSE1 = sum(small.model$residuals^2)
  SSE2 = sum(big.model$residuals^2)
  PR2 = (SSE1 - SSE2) / SSE1
  return(PR2)
}

# Give me multipliers
find.mult = function(alpha,a,b,dfSSE,g,group){
  if(group == "A"){
    Tuk = round(qtukey(1-alpha,a,dfSSE)/sqrt(2),3)
    Bon = round(qt(1-alpha/(2*g), dfSSE ) ,3)
    Sch = round(sqrt((a-1)*qf(1-alpha, a-1, dfSSE)),3)
  }else if(group == "B"){
    Tuk = round(qtukey(1-alpha,b,dfSSE)/sqrt(2),3)
    Bon = round(qt(1-alpha/(2*g), dfSSE ) ,3)
    Sch = round(sqrt((b-1)*qf(1-alpha, b-1, dfSSE)),3)
  }else if(group == "AB"){
    Tuk = round(qtukey(1-alpha,a*b,dfSSE)/sqrt(2),3)
    Bon = round(qt(1-alpha/(2*g), dfSSE ) ,3)
    Sch = round(sqrt((a*b-1)*qf(1-alpha, a*b-1, dfSSE)),3)
  }
  results = c(Bon, Tuk,Sch)
  names(results) = c("Bonferroni","Tukey","Scheffe")
  return(results)
}

# Give me CI
give.me.CI = function(data,MSE,equal.weights = TRUE,multiplier,group,cs){
   if(sum(cs) != 0 & sum(cs !=0 ) != 1){
    return("Error - you did not input a valid contrast")
  }else{
    the.means = find.means(data)
    the.ns =find.means(data,length)
    nt = nrow(data)
    a = length(unique(data[,2]))
    b = length(unique(data[,3]))
    if(group =="A"){
      if(equal.weights == TRUE){
        a.means = rowMeans(the.means$AB)
        est = sum(a.means*cs)
        mul = rowSums(1/the.ns$AB)
        SE = sqrt(MSE/b^2 * (sum(cs^2*mul)))
        N = names(a.means)[cs!=0]
        CS = paste("(",cs[cs!=0],")",sep = "")
        fancy = paste(paste(CS,N,sep =""),collapse = "+")
        names(est) = fancy
      } else{
        a.means = the.means$A
        est = sum(a.means*cs)
        SE = sqrt(MSE*sum(cs^2*(1/the.ns$A)))
        N = names(a.means)[cs!=0]
        CS = paste("(",cs[cs!=0],")",sep = "")
        fancy = paste(paste(CS,N,sep =""),collapse = "+")
        names(est) = fancy
      }
    }else if(group == "B"){
      if(equal.weights == TRUE){
        b.means = colMeans(the.means$AB)
        est = sum(b.means*cs)
        mul = colSums(1/the.ns$AB)
        SE = sqrt(MSE/a^2 * (sum(cs^2*mul)))
        N = names(b.means)[cs!=0]
        CS = paste("(",cs[cs!=0],")",sep = "")
        fancy = paste(paste(CS,N,sep =""),collapse = "+")
        names(est) = fancy
      } else{
        b.means = the.means$B
        est = sum(b.means*cs)
        SE = sqrt(MSE*sum(cs^2*(1/the.ns$B)))
        N = names(b.means)[cs!=0]
        CS = paste("(",cs[cs!=0],")",sep = "")
        fancy = paste(paste(CS,N,sep =""),collapse = "+")
        names(est) = fancy
      }
    } else if(group == "AB"){
      est = sum(cs*the.means$AB)
      SE = sqrt(MSE*sum(cs^2/the.ns$AB))
      names(est) = "someAB"
    }
    the.CI = est + c(-1,1)*multiplier*SE
    results = c(est,the.CI)
    names(results) = c(names(est),"lower bound","upper bound")
    return(results)
  }
}
```

# Information for Finding Model
$\alpha = `r alpha`$

Model with interaction:

$Y_{ijk} = \mu_{..} + \gamma_i + \delta_j + (\gamma\delta)_{ij} + \epsilon_{ijk}$

Model with factor A and factor B:

$Y_{ijk} = \mu_{..} + \gamma_i + \delta_j + \epsilon_{ijk}$

Model with factor A only:

$Y_{ijk} = \mu_{..} + \gamma_i + \epsilon_{ijk}$

Model with factor B only:

$Y_{ijk} = \mu_{..} + \delta_j + \epsilon_{ijk}$

Final model after conducting F-statistic tests:

$Y_{ijk} = \mu_{..} + \gamma_i + \delta_j + \epsilon_{ijk}$

With constraints: $\sum \gamma_i = 0$, $\sum \delta_j = 0$

$H_0:$ Use full model

$H_a:$ Use reduced model

# Model Selection Table
```{r Model Selection Table}
model_selection <- data.frame(
  "Model" = c(
    "Interaction Effect",
    "Factor A Effect (Region)",
    "Factor B Effect (House Size)"
  )
)

model_selection$"$R^2$" <- c(
  Partial.R2(A.B, AB),
  Partial.R2(B, A.B),
  Partial.R2(A, A.B)
)

model_selection$"$F_s$" <- c(
  anova(A.B, AB)$F[2],
  anova(B, A.B)$F[2],
  anova(A, A.B)$F[2]
)

model_selection$"$p$" <- c(
  anova(A.B, AB)$"Pr(>F)"[2],
  anova(B, A.B)$"Pr(>F)"[2],
  anova(A, A.B)$"Pr(>F)"[2]
)

model_selection$"$p \\leq \\alpha$" <- model_selection$"$p$" <= alpha

model_selection$Include <- model_selection$"$p \\leq \\alpha$"

knitr::kable(t(model_selection))
```

# Analysis - Pairwise CIs for Factor A
```{r Get SSE}
all.models = list(AB,A.B,A,B,N)
SSE = t(as.matrix(sapply(all.models,function(M) sum(M$residuals^2))))
colnames(SSE) = c("AB","(A+B)","A","B","Empty/Null")
rownames(SSE) = "SSE"
```

```{r Analysis Setup}
# Get relavent values for CI based on model choice
n_T <- nrow(data)
a <- length(unique(data$A))
b <- length(unique(data$B))
sse <- SSE["SSE", "(A+B)"]
df_sse <- n_T - a - b + 1
mse <- sse / df_sse
```

```{r Get Multiplier for Factor A}
all.mult <- find.mult(alpha = alpha, a = a, b = b, dfSSE = df_sse, g = 6,
                      group = "A")
the.mult <- min(all.mult)
knitr::kable(all.mult)
```

We can pick Bonferroni, Tukey, or Scheffe multiplier. We pick the smallest multiplier, Bonferroni $(`r the.mult`)$, for more precision.

```{r Do Confidence Intervals for Factor A}
# Get combinations of all potential confidence intervals
CIs.A <- data.frame(Values = c("Mean", "Lower", "Upper"))
CIs.A$"$\\mu_{1.} - \\mu_{6.}$" <- give.me.CI(
  data, mse, equal.weights = TRUE,
  the.mult, "A", c(1, 0, 0, 0, 0, -1)
)
CIs.A$"$\\mu_{2.} - \\mu_{1.}$" <- give.me.CI(
  data, mse, equal.weights = TRUE,
  the.mult, "A", c(-1, 1, 0, 0, 0, 0)
)

CIs.A$"$\\mu_{5.} - \\mu_{2.}$" <- give.me.CI(
  data, mse, equal.weights = TRUE,
  the.mult, "A", c(0, -1, 0, 0, 1, 0)
)

CIs.A$"$\\mu_{3.} - \\mu_{5.}$" <- give.me.CI(
  data, mse, equal.weights = TRUE,
  the.mult, "A", c(0, 0, 1, 0, -1, 0)
)

CIs.A$"$\\mu_{4.} - \\mu_{3.}$" <- give.me.CI(
  data, mse, equal.weights = TRUE,
  the.mult, "A", c(0, 0, -1, 1, 0, 0)
)

CIs.A$"$\\mu_{4.} - \\mu_{6.}$" <- give.me.CI(
  data, mse, equal.weights = TRUE,
  the.mult, "A", c(0, 0, 0, 1, 0, -1)
)

CIs.A <- as.data.frame(t(CIs.A))
colnames(CIs.A) <- CIs.A[1, ]
CIs.A <- CIs.A[-1, ]

CIs.A$Significant <- FALSE

knitr::kable(CIs.A)
```

# Analysis - Pairwise CIs for Factor B
```{r Get Multiplier for Factor B}
all.mult <- find.mult(alpha = alpha, a = a, b = b, dfSSE = df_sse, g = 4,
                      group = "B")
the.mult <- min(all.mult)
knitr::kable(all.mult)
```

We can pick Bonferroni, Tukey, or Scheffe multiplier. We pick the smallest multiplier, Bonferroni $(`r the.mult`)$, for more precision.

```{r Do Confidence Intervals for Factor B}
# Get combinations of all potential confidence intervals
CIs.B <- data.frame(Values = c("Mean", "Lower", "Upper"))

CIs.B$"$\\mu_{.2} - \\mu_{.4}$" <- give.me.CI(
  data, mse, equal.weights = TRUE,
  the.mult, "B", c(0, 1, 0, -1)
)

CIs.B$"$\\mu_{.1} - \\mu_{.2}$" <- give.me.CI(
  data, mse, equal.weights = TRUE,
  the.mult, "B", c(1, -1, 0, 0)
)

CIs.B$"$\\mu_{.3} - \\mu_{.1}$" <- give.me.CI(
  data, mse, equal.weights = TRUE,
  the.mult, "B", c(-1, 0, 1, 0)
)

CIs.B$"$\\mu_{.3} - \\mu_{.4}$" <- give.me.CI(
  data, mse, equal.weights = TRUE,
  the.mult, "B", c(0, 0, 1, -1)
)

CIs.B <- as.data.frame(t(CIs.B))
colnames(CIs.B) <- CIs.B[1, ]
CIs.B <- CIs.B[-1, ]

CIs.B$Significant <- FALSE

knitr::kable(CIs.B)
```

# Appendix
```{r appendix, results="asis", echo = FALSE}
# Get all the labels
all_labels <- knitr::all_labels()

# Print out code along with labels
for (label in all_labels) {
  if (label == "setup" || label == "appendix") {
    next
  }

  code <- knitr::knit_code$get(label)

  cat("```r \n")
  cat("#", label, "\n")
  cat(code, sep = "\n")
  cat("\n```\n")
}
```