---
title: "Your Title"
author: "Your Name"
---

<!-- Uncomment this part if you want to hide code in output -->
<!-- ```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
``` -->

# Basics
Make sure to categorize chunks of code under headers for more organization. 

```{r Chunk1}
# This is a chunk of code, name is Chunk1
# You may not have duplicate chunk names
# You may name your chunks anything other than setup or appendix
x <- 1
x # You can output variables like this
```

```{r Chunk2}
y <- 2
```

You can also output variables ($y = `r y`$) inline.

```{r Chunk3}
z <- 3.1415
round(z, 2) # You can also round large numbers
```

# Formatting Nice Dataframes
```{r DataFrames}
the.data <- data.frame(y = c(1, 2, 3, 4), x = c(1, 2, 3, 4))
the.data # We lose points if we format like this in final report
```

```{r DataFramesNicelyFormatted}
knitr::kable(the.data) # You need to format like this
```

# Converting Matrices to Nice Tables
```{r FormatMatrixNicely}
the.matrix <- matrix(c(1, 2, 3, 4, 5, 6), nrow = 3, ncol = 2)
knitr::kable(data.frame(the.matrix))
```

# Extract information from models
```{r ExtractInfoFromModel}
x <- c(1, 2, 3, 4, 5, 6, 7, 8)
y <- runif(8) * 10 # This vector is random every time
the.model <- lm(y ~ x) # This can be any model
the.summary <- summary(the.model)
the.summary
```

- Extracted p-value of intercept: $p = `r the.summary$coefficients["(Intercept)", "Pr(>|t|)"]`$
- You can use chatGPT to figure out how to extract values from `the.summary`

# Need Help Writing R Code?
- Consult the R examples that professor put on piazza
- If you cannot find the answer in provided R examples:
   - Ask ChatGPT, but make sure you understand what's happening in the output
   - Conduct a Google search

# Appendix
After converting to word document, you will need to copy all the code here to word document when done.

```{r appendix, results="asis", echo = FALSE}
# Get all the labels
all_labels <- knitr::all_labels()

# Print out code along with labels
for (label in all_labels) {
  if (label == "setup" || label == "appendix") {
    next
  }

  code <- knitr::knit_code$get(label)

  cat("```r \n")
  cat("# ", label, "\n")
  cat(code, sep = "\n")
  cat("\n```\n")
}
```